---
title: "Data Preparation and Cleaning"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(mltools)
library(ggplot2)
library(data.table)
```

#### Loading Data

```{r}
spotify.df.full = read.csv("data/SpotifyFeatures.csv")

```

#### Basic EDA

```{r}
print("Dimension")
print(dim(spotify.df.full))

# Renaming genre column
colnames(spotify.df.full)[colnames(spotify.df.full) == "ï..genre"] <- "genre"

print("Columns")
print(colnames(spotify.df.full))
```

#### Summary 

```{r}
summary(spotify.df.full)
```

Dependent Variable: popularity
Continouos Variables: Valence, Tempo, Speechiness, Loudness, Liveness, Intrumentalness, Energy, Duration, Dancibility, Acousticness (10)
Categorical Variables: Time Signature, Mode, Key, Genre (10)
Unique Identifier: Track ID
Other Variables:Artist Name, Track Name

```{r}
# Categorical and Continuous variables list
categorical.vars = c("time_signature", "mode", "key", "genre")
print(categorical.vars)
unique.identifier = c("track_id")
print(unique.identifier)
other.vars = c("track_name", "artist_name")
print(other.vars)
dependent.var = c("popularity")
print(dependent.var)
continuous.vars = c("valence", "tempo", "speechiness", "loudness", "liveness", "instrumentalness", "energy", "duration_ms", "danceability", "acousticness")
print(continuous.vars)
```


#### Distinct Values for categorical variables

```{r}
print("Genre")
print(length(unique(spotify.df.full$genre)))
table(spotify.df.full$genre)

print("Time Signature")
table(spotify.df.full$time_signature)

print("Mode")
table(spotify.df.full$mode)

print("Key")
print(length(unique(spotify.df.full$key)))
table(spotify.df.full$key)
```


#### Checking number of missing values
```{r}
colSums(is.na(spotify.df.full))
```

No missing values in the dataset

#### Outlier Treatment

```{r echo=FALSE}
# Checking number of outliers for continuous variables
outliers = vector()
for (c in continuous.vars) {
  o = length(boxplot(spotify.df.full$c, plot = FALSE)$out)
  outliers = c(outliers, o)
}

outlier.df = data.frame(
  column.name = continuous.vars,
  no.of.outliers = outliers
)

print(outlier.df)

```

Therefore, outlier treament is not needed as number of outliers is 0 for all the continuous variables.

#### Standardize Continuous variables

```{r}
cont.vars = colnames(spotify.df.full)[colnames(spotify.df.full) %in% continuous.vars]
spotify.df.full.cont = spotify.df.full[, continuous.vars]
spotify.df.full.cont.scaled = data.frame(scale(spotify.df.full.cont))
```



#### One-Hot Encoding

```{r}
cat.vars = colnames(spotify.df.full)[colnames(spotify.df.full) %in% categorical.vars]
spotify.df.full.cat = spotify.df.full[, cat.vars]
spotify.df.full.cat = data.table(spotify.df.full.cat)
spotify.df.full.cat.onehot = one_hot(spotify.df.full.cat, cols = "auto")
```

#### Other Variables

```{r}
other.vars = colnames(spotify.df.full)[colnames(spotify.df.full) %in% other.vars]
spotify.df.full.other = spotify.df.full[, other.vars]
```

#### Final Dataset

```{r}
spotify.final.df = cbind(spotify.df.full$track_id, spotify.df.full.other, spotify.df.full.cont.scaled, spotify.df.full.cat.onehot,spotify.df.full$popularity)
colnames(spotify.final.df) = c("track_id", colnames(spotify.df.full)[2:59], "popularity")
```

```{r}
# Write dataset
write.csv(spotify.final.df, "data/spotify.dataset.final.csv")
```





________________________________________________________________________________________________________________
To Do-
  1.  Missing Value Treatment
  2.  Outlier Treatment
  3.  Standardize continuous variables
  4.  One-hot encode categorical variables
