---
title: "Linear Regression"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(caTools)
```

#### Loading Data

```{r}
spotify.final.df = read.csv("data/spotify.dataset.final.csv")
spotify.final.df = spotify.final.df[, !colnames(spotify.final.df) %in% "X"]
```

```{r}
print(dim(spotify.final.df))
print(colnames(spotify.final.df))
```

```{r}
# Create a dataset for building models
non.model.vars = c("track_id", "artist_name", "track_name")
spotify.model.df = spotify.final.df[, !colnames(spotify.final.df) %in% non.model.vars]
```

```{r}
# Dividing data into test and train
set.seed(123)
sample = sample.split(spotify.model.df, SplitRatio = 0.75)
spotify.model.df.train = subset(spotify.df.model, sample == TRUE)
spotify.model.df.test = subset(spotify.df.model, sample == FALSE)
```

#### Fitting a linear model

```{r}
# Fitting a linear model
fit1 = lm(popularity ~ ., data = spotify.model.df.train)
summary(fit1)
```

Few coefficients are not defined, variables - time_signature_5.4, mode_Minor, key_G., genre_World 
Few of the key, time signature variables are not significant. genre_World is also not significant

```{r}
# Checking correlation of "time_signature_5.4" variable with other variables
cor.time_signature_5.4 = cor(spotify.model.df[, !colnames(spotify.model.df) %in% "time_signature_5.4"], spotify.model.df$time_signature_5.4)
print(cor.time_signature_5.4[cor.time_signature_5.4 %in% cor.time_signature_5.4[abs(cor.time_signature_5.4) > 0.3]])
```

```{r}
# Checking correlation of "mode_Minor" variable with other variables
cor.mode_Minor = cor(spotify.model.df[, !colnames(spotify.model.df) %in% "mode_Minor"], spotify.model.df$mode_Minor)
print(cor.mode_Minor[cor.mode_Minor %in% cor.mode_Minor[abs(cor.mode_Minor) > 0.3]])

```

```{r}
names(fit1)
```

#### Model Evaluation on test data

```{r}
# Model Evaluation on test data
test.pred.fit1 = predict(fit1, spotify.df.model.test)

print("RMSE")
print(rmse(test.pred.fit1, spotify.df.model.test$popularity))

rsquared <- function(predicted.vals, actual.val) {
  SS.total = sum((actual.val - mean(actual.val))^2)
  SS.residual = sum((actual.val - predicted.vals)^2)
  
  test.r2 = 1 - (SS.residual/SS.total)
  return(test.r2)
}

print("R-Squared")
print(rsquared(test.pred.fit1, spotify.df.model.test$popularity))

```

```{r}
plot(fit1)
```

```{r}
# Create a dataset without 'genre' variables
spotify.df.model.train.wo.genre = spotify.df.model.train[,-grep("genre", colnames(spotify.df.model.train))]
spotify.df.model.test.wo.genre = spotify.df.model.test[,-grep("genre", colnames(spotify.df.model.test))]
```

#### Fitting a linear model without genre variable

```{r}
# Fitting a linear model without genre variable
fit2 = lm(popularity ~ ., data = spotify.df.model.train.wo.genre)
summary(fit2)
```

Low R-Squared value, therefore, no further investigation will be performed.

#### Fitting a linear model with interaction with duration

```{r}
fit3 = lm(popularity~.*duration_ms, data = spotify.df.model.train)
summary(fit3)
```

Not much improvement in R-Squared value, therefore, no further investigation will be performed.

